D-Hammer: Efficient Equational Reasoning for Labelled Dirac Notation

### 2025-04-05 06related_conclusion.tex
.

\paragraph*{Comparison with ZX Calculus}
The ZX calculus~\cite{DBLP:conf/icalp/CoeckeD08,vandewetering2020zx}
is a popular graphical calculus for reasoning about quantum states.
The foundations of the ZX calculus are grounded in categorical quantum
mechanics (CQM)~\cite{DBLP:conf/lics/AbramskyC04}, which provides an
category-theoretic foundation for quantum mechanics.  A key aspect of
the ZX calculus and related graphical calculi is that they can be
equipped with powerful rewriting systems. These systems support a form
of graph rewriting, i.e.\, they exploit the graph representation of
quantum states. Two notable properties of these systems are:
\begin{itemize}
\item completeness:
\item efficiency:
\end{itemize}
Both Labelled Dirac Notation (LDN) and the ZX calculus share the use
of rewriting techniques. However, it is difficult to compare the ZX
calculus and LDN. First, they have different scopes: the ZX calculus
mainly focuses on quantum circuits, whereas LDN aims to provide
general support for quantum physics. This difference in scope is
reflected in the underlying languages: for instance, many works in ZX
calculus consider concrete sets of quantum gates, whereas LDN
emphasizes the use of big operators, and does not attempt to provide
specialized support for a particular set of quantum gates. As a
result, many of the examples handled by D-Hammer cannot be formalized
in existing ZX-calculus tools. On the contrary, D-Hammer is not
competitive on the examples that can be verified by ZX-calculus.

>>>>

\paragraph*{Comparison with ZX Calculus}
The ZX calculus~\cite{DBLP:conf/icalp/CoeckeD08,vandewetering2020zx}
is a popular graphical calculus for reasoning about quantum states.
The foundations of the ZX calculus are grounded in categorical quantum
mechanics (CQM)~\cite{DBLP:conf/lics/AbramskyC04}, which provides an
category-theoretic foundation for quantum mechanics.  A key aspect of
the ZX calculus and related graphical calculi is that they support a
form of graph rewriting. Two notable properties of these rewriting
systems are:
\begin{itemize}
\item completeness:
\item efficiency:
\end{itemize}
Both Labelled Dirac Notation (LDN) and the ZX calculus share the use
of rewriting techniques. However, it is difficult to compare the ZX
calculus and LDN. First, they have different scopes: the ZX calculus
mainly focuses on quantum circuits, whereas LDN aims to provide
general support for quantum physics. This difference in scope is
reflected in the underlying languages: for instance, many works in ZX
calculus consider concrete sets of quantum gates, whereas LDN
emphasizes the use of big operators, and does not attempt to provide
specialized support for a particular set of quantum gates. As a
result, many of the examples handled by D-Hammer cannot be formalized
in existing ZX-calculus tools. On the contrary, D-Hammer is not
competitive on the examples that can be verified by ZX-calculus.
